---
title: "Reproducible Research: Peer Assessment 1"
author: "Walter Parada L"
date: "12/11/2020"
output: html_document
Keep_md: true

---
## Set Working Directory  

```{r setworkingdirectory, echo=TRUE}
setwd("C:/WALTER/Capacitacion/DATA SCIENCE/MODULO 5 - Reproducible Research/RepData_PeerAssessment1")
```

## Loading and preprocessing the data

```{r readdata}
activity<-read.csv(file="activity.csv")
head(activity)
names(activity)
```
### Database structure
```{r}
str(activity)
```
### convert var date chr to date 
```{r addvarDate}
activity$date<-as.Date(x=activity$date,format ="%Y-%m-%d")
str(activity)
```


## What is mean total number of steps taken per day?

```{r Mean Total Steps}
steps <- aggregate(steps~date,activity,sum, na.rm = TRUE)
mean <- mean(steps$steps)
print(sprintf("Mean total of steps per day are: %f",mean))
```
```{r}
steps <- aggregate(steps~date,activity,sum, na.rm = TRUE)
hist(steps$steps,main = "Total number of steps taken per day and mean", xlab = "TOTAL STEPS/DAY", ylab= "Days", col = "red")
abline(v = mean, col = "blue", lwd = 2)
```
  
**Most of days are between 10000 to 15000**  


## What is the average daily activity pattern?

```{r}
average<- aggregate(activity$steps,by = list(activity$interval), mean,na.rm = TRUE)
names(average) <- c("interval", "steps")
plot(x = average$interval,y = average$steps,type = "l",xlab = "5 Minute Interval", ylab = "Average Number of Steps",main = "Average Daily Activity Planner")
```
```{r Max Avg Daily Number of Steps}
max.average<- average$interval[average$steps == max(average$steps)]
print(sprintf("5-min time Interval with maximum average steps taken per day: %i",max.average))
```
## Imputing missing values

Calculate missing values cases

```{r Missing Values}
na.count <- sum(!complete.cases(activity))
print(sprintf("Total missing Values in the data is: %i", na.count))

```
```{r}
## unique interval in activity
uniqueint<- unique(activity$interval)
## make a copy of dataset activity
copydata <- activity

## Strategy is fill missing values with mean for that 5-minute interval

MissingData <- function(timeint) {
    copydata[which(activity$interval==timeint & is.na(activity$steps)),
             ]$steps <<- mean(activity[which(activity$interval==timeint & !is.na(activity$steps)), ]$steps)
}

lapply(uniqueint, MissingData)
```
```{r}
Newsteps <- aggregate(steps~date, copydata, sum, na.rm=TRUE)
hist(Newsteps$steps, 
     main="Mean total steps taken per day (with Missing values imputed)", 
     xlab="Steps", 
     ylab="Number of Days", 
     ylim=c(0,40),
     col="green")
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
## Create a column with Weekday or Weekend :
copydata$type <- weekdays(as.Date(copydata$date))
copydata$type[copydata$type %in% c('sabado','domingo')] <-"Weekend"
copydata$type[copydata$type != "Weekend"] <-"Weekday"
copydata$type <- as.factor(copydata$type)

Weekday_data <- subset(copydata, type=="Weekday")
Weekday_stepsToInterval<- aggregate(Weekday_data$steps,by = list(Weekday_data$interval), mean,na.rm = TRUE)
names(Weekday_stepsToInterval) <- c("interval", "steps")

Weekend_data <- subset(copydata, type=="Weekend")
Weekend_stepsToInterval<- aggregate(Weekend_data$steps,by = list(Weekend_data$interval), mean,na.rm = TRUE)
names(Weekend_stepsToInterval) <- c("interval", "steps")

## Create two line plots:
par(mfrow=c(2,1), mar=c(4,4,2,1), oma=c(0,0,2,0))

with(copydata, {
    plot(steps~interval,data=Weekday_stepsToInterval,type="l",
main="Steps for Weekdays", xlab="Interval", col="red")
    plot(steps~interval,data=Weekend_stepsToInterval,type="l",
main="Steps for Weekends", xlab="Interval", col="red")
    })
```

+ People walk more on weekends specially after 10:00  
